---
title: "Exploratory Data Analysis - Loan Data"
output:
  html_document: html_document
  html_notebook: html_notebook
---


# Packages requiered

These packages are requiered for this EDA. To install, please execute the following code.

*install.packages("knitr", dependencies = T)*
*install.packages("dplyr", dependencies = T)*
*install.packages("ggplot2", dependencies = T)*
*install.packages("RColorBrewer", dependencies = T)* 
*install.packages('ggthemes', dependencies = TRUE)* 


```{r}
library("knitr")
library("dplyr")
library("ggplot2")
library("RColorBrewer")
library("ggthemes")
```

# Set directory path

```{r}
getwd()
setwd('/Users/aguijarro/SoftwareCode/github.com/aguijarro/R/Project')
getwd()
list.files()
```

# Load data

```{r}
loanData <- read.csv('prosperLoanData.csv')
```

# Get dim data

```{r}
dim(loanData)
```

# Explore some data

```{r}
head(loanData,5)
tail(loanData,5)
```

# Get column types

```{r}
lapply(loanData, typeof)

```

# Display the estructure of the object

```{r}
str(loanData)
summary(loanData)
```

# Notes

When we use typeof function to get the data type of column, the result in some cases are integer, meanwhile, using str function the result is a Factor.

For example: 

typeof()
$ListingCreationDate
[1] "integer"

str()
$ ListingCreationDate                : Factor w/ 113064 levels 

# Analysis Date Variables

Transform to Date
-----------------
```{r}
ListingCreationDate<-as.Date(loanData$ListingCreationDate)
```

Get and Create Listing Creation Date - Years
---------------------------------------------
```{r}
ListingCreationYear<-format(ListingCreationDate, "%Y")
loanData$ListingCreationYear <- as.factor(as.numeric(ListingCreationYear))
```

Get and Create Listing Creation Date - Months
---------------------------------------------
```{r}
ListingCreationMonth<-format(ListingCreationDate, "%m")
loanData$ListingCreationMonth <- as.factor(as.numeric(ListingCreationMonth))
```

Get and Create Listing Creation Date - Days
---------------------------------------------
```{r}
ListingCreationDay<-format(ListingCreationDate, "%d")
loanData$ListingCreationDay <- as.factor(as.numeric(ListingCreationDay))
```



Graph Listing Creation Date - Days
-----------------------------------

```{r}
qplot(data = loanData, x = ListingCreationDay) + 
  scale_x_discrete(breaks=1:31)
```

```{r}
qplot(data = loanData, x = ListingCreationDay) + 
  scale_x_discrete(breaks=01:31) +
  facet_wrap(~ListingCreationMonth, ncol=3)

```



Graph Listing Creation Date - Months
-----------------------------------

```{r}
qplot(data = loanData, x = ListingCreationMonth) + 
  scale_x_discrete(breaks=1:12)
```

```{r}
qplot(data = loanData, x = ListingCreationMonth) + 
  scale_x_discrete(breaks=1:12) +
  facet_wrap(~ListingCreationYear, ncol=3)
```


Graph Listing Creation Date - Year
-----------------------------------

```{r}
qplot(data = loanData, x = ListingCreationYear)
```





# Basic analysis

```{r}

```

# Analyced factor variables

```{r}
table(loanData$EmploymentStatus)
levels(loanData$EmploymentStatus)
```

# Graph factor variables

```{r}
loanDataOrder <- ordered(loanData$EmploymentStatus, levels = c("Employed", "Full-time", "Not available", "Not employed", "Other", "Part-time", "Retired", "Self-employed"))
qplot(data = loanData, x = loanDataOrder) + 
  scale_x_discrete(breaks=1:31)
```

# Analyzing ListingCreationDate variable
```{r}

loanData$ListingCreationDate

```